---
title: "serving_dog_training_Rmd"
output: html_document
date: "2024-10-09"
---

First of all, loading the data
```{r}
dog_training <- read.csv('/Users/ashly/Downloads/dog_training_data.csv')

head(dog_training)
```

Check some information via Statistical Approach

```{r}
summary(dog_training)
```

Scatter Plot with Regression Lines

```{r}

library(ggplot2)

data <- data.frame(
  Dog_Breed = c('Golden Retriever', 'Border Collie', 'Poodle', 'Beagle', 'Doberman', 
                'German Shepherd', 'Husky', 'Labrador', 'Rottweiler', 'Belgian Malinois'),
  Training_Hours = c(285.71, 175.56, 303.48, 206.86, 345.68, 233.43, 180.71, 314.89, 398.41, 343.59),
  Task_Completion_Rate = c(0.33, 0.2, 0.33, 0.27, 0.4, 0.27, 0.2, 0.33, 0.33, 0.4)
)


ggplot(data, aes(x=Training_Hours, y=Task_Completion_Rate, color=Dog_Breed)) +
  geom_point(size=3) +  # Scatter plot
  geom_smooth(method="lm", se=FALSE) +  # Add linear regression lines
  labs(title="Impact of Dog Breed and Training Hours on Task Completion Rate",
       x="Training Hours",
       y="Task Completion Rate") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

```
Regression Model

```{r}

data$Dog_Breed <- as.factor(data$Dog_Breed)


model <- lm(Task_Completion_Rate ~ Training_Hours + Dog_Breed, data=data)
model

summary(model)

```



Analysis the Dog Breed and Health Score with some visualization

```{r}
library(ggplot2)

# Dog Breed
breed_distribution <- as.data.frame(table(dog_training$Dog_Breed))
colnames(breed_distribution) <- c("Breed", "Count")
breed_distribution$Percentage <- round((breed_distribution$Count / sum(breed_distribution$Count)) * 100, 1)

# visualization
ggplot(breed_distribution, aes(x = "", y = Count, fill = Breed)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar(theta = "y") +
  theme_void() +
  ggtitle("Dog Breed Distribution") +
  geom_text(aes(label = paste0(Percentage, "%")), position = position_stack(vjust = 0.5))

# Health Score
dog_training$Health_Score_Cut <- cut(dog_training$Health_Score, breaks = c(60, 70, 80, 90, 100))
health_score_distribution <- as.data.frame(table(dog_training$Health_Score_Cut))
colnames(health_score_distribution) <- c("Health_Score_Range", "Count")
health_score_distribution$Percentage <- round((health_score_distribution$Count / sum(health_score_distribution$Count)) * 100, 1)

# visualization
ggplot(health_score_distribution, aes(x = "", y = Count, fill = Health_Score_Range)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar(theta = "y") +
  theme_void() +
  ggtitle("Health Score Distribution") +
  geom_text(aes(label = paste0(Percentage, "%")), position = position_stack(vjust = 0.5))


```


```{r}
# If I want to get the Top3 breeds of dogs' count and percentage 
library(dplyr)

breed_distribution <- as.data.frame(table(dog_training$Dog_Breed))
colnames(breed_distribution) <- c("Breed", "Count")

breed_distribution$Percentage <- round(breed_distribution$Count/sum(breed_distribution$Count)*100, 1)

top_3_breeds <- breed_distribution %>% 
  arrange(desc(Count)) %>%
  head(3)

top_3_breeds

```


```{r}
ggplot(dog_training, aes(x = Training_Hours, y = Task_Completion_Rate)) +
  geom_point() +
  facet_grid(Dog_Breed ~ Health_Score_Cut) +
  ggtitle("Task Completion Rate by Breed and Health Score Range") +
  xlab("Training Hours") +
  ylab("Task Completion Rate") +
  theme(
    strip.text.y = element_text(size = 8, angle = 0),  
    strip.text.x = element_text(size = 8),             
    axis.text = element_text(size = 8),                
    axis.title = element_text(size = 8),               
    plot.title = element_text(size = 12)              
  )

```

Decision tree
Step1: label the top 10% fastest learners based on Task Completion Rate and Training Hours
```{r}

#install.packages("caret")
#install.packages("rpart.plot")

library(dplyr)
library(caret)
library(rpart)
library(rpart.plot)

# Create a 'Fast_Learner' label by ranking dogs based on Task Completion Rate and Training Hours
# Assuming faster learners have high task completion rate and low training hours
dog_training <- dog_training %>%
  mutate(Learning_Score = Task_Completion_Rate / Training_Hours) %>%
  arrange(desc(Learning_Score)) %>%
  mutate(Fast_Learner = ifelse(row_number() <= 0.1 * n(), 1, 0))  # Top 10% as Fast Learners

table(dog_training$Fast_Learner)


```

Step2: Split the data into training set and testing set
```{r}

library(dplyr)
library(caret)
library(rpart)
library(rpart.plot)

dog_training <- read.csv('/Users/ashly/Downloads/dog_training_data.csv')

set.seed(123)

# split the data into training (80%) and testing (20%) sets
sample_index <- sample(1:nrow(dog_training), size = 0.8 * nrow(dog_training))

training_set <- dog_training[sample_index, ]
testing_set <- dog_training[-sample_index, ]

training_set
testing_set

```

Step3: Training the decision tree model via rpart package
```{r}
# Step1
dog_training <- dog_training %>%
  mutate(Learning_Score = Task_Completion_Rate / Training_Hours) %>%
  arrange(desc(Learning_Score)) %>%
  mutate(Fast_Learner = ifelse(row_number() <= 0.1 * n(), 1, 0))  

table(dog_training$Fast_Learner)



# Step2
set.seed(123)

split_index <- createDataPartition(dog_training$Fast_Learner, p = 0.8, list = FALSE)
training_set <- dog_training[split_index, ]
testing_set <- dog_training[-split_index, ]



# Step3
train_control <- trainControl(method = "cv", number = 10)
model <- train(Fast_Learner ~ Training_Hours + Task_Completion_Rate + Health_Score + Stress_Level + Dog_Breed,
                     data = training_set,
                     method = "rpart",
                     trControl = train_control,
                     tuneLength = 10)

# View the model parameters
model$bestTune


rpart.plot(model$finalModel, type = 3, extra = 101, fallen.leaves = TRUE, main = "Decision Tree for Fast Learners")



```



Step4: Evaluating the Model
```{r}

predictions <- predict(decision_tree_model, testing_set, type = "class")

# Create a confusion matrix to evaluate the performance
confusionMatrix(as.factor(predictions), as.factor(testing_set$Fast_Learner))

```
